Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2016-05-06 doc/invt.tex
small edits

\documentclass[reprint, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}

@@

Free energy calculation\cite{frenkel, newman} is a central theme
in computational physics and chemistry.
%
Given a system,
the task is to compute,
via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
and molecular dynamics\cite{frenkel, karplus2002} (MD) simulations,
a distribution, $p^*(z)$,
along a collective variable $z$.
%
However, $z$
often represents a slow variable
with a localized or multi-modal distribution, $p^*(z)$.
%
Thus,
to capture the global shape of the free energy surface
in terms of the potential of mean force (PMF), $-\ln p^*(z)$,
an effective strategy is to sample a flat distribution
along $z$ by introducing a bias potential that
cancels the PMF.
%The negative logarithm,
%$-\ln p^*(z)$,
%defines a free energy,
%or the potential of mean force (PMF).
%%
%Normally, the distribution $p^*(z)$ is
%limited in one or a few narrow $z$ regions.
%%
%Therefore, to capture the global shape
%of the PMF,
%one has to run multiple independent simulations
%under different umbrella or bias potentials
%along $z$\cite{mezei1987, berg1992, lee1993},
%and then piece together the entire curve.
%%
%This solution can be laborious % clumsy in practice
%and it often fails to converge quickly
%for a complex and glassy system,
%because the system can be trapped
%in a local free energy minimum.
%%for a long time.
%
%
%
%A better solution is to artificially flatten
%the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
%wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
%%
%This alleviates the above problem of local trapping
%by forcing the system to travel more
%deliberately along $z$,
%which often represents
%a direction of slow motion.
%%
%To achieve a flat distribution, however,
%the method must be able to construct a
%bias potential that cancels the PMF.

>>>>

\documentclass[preprint, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{upgreek}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{hyperref}

@@

Free energy calculation\cite{frenkel, newman} is a central theme
in computational physics and chemistry.
%
Given a system,
the task is to compute,
via either Monte Carlo\cite{frenkel, newman, landau_binder} (MC)
and molecular dynamics\cite{frenkel, karplus2002} (MD) simulations,
a distribution, $p^*(z)$,
along a collective variable $z$.
%
However, $z$
often represents a slow variable
with a localized or multi-modal distribution, $p^*(z)$.
%
Thus,
to capture the global shape of the free energy surface
in terms of the potential of mean force (PMF), $-\ln p^*(z)$,
an effective strategy is to sample a flat distribution
along $z$ by introducing a bias potential that
cancels the PMF\cite{mezei1987, berg1992, lee1993,
wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
%The negative logarithm,
%$-\ln p^*(z)$,
%defines a free energy,
%or the potential of mean force (PMF).
%%
%Normally, the distribution $p^*(z)$ is
%limited in one or a few narrow $z$ regions.
%%
%Therefore, to capture the global shape
%of the PMF,
%one has to run multiple independent simulations
%under different umbrella or bias potentials
%along $z$\cite{mezei1987, berg1992, lee1993},
%and then piece together the entire curve.
%%
%This solution can be laborious % clumsy in practice
%and it often fails to converge quickly
%for a complex and glassy system,
%because the system can be trapped
%in a local free energy minimum.
%%for a long time.
%
%
%
%A better solution is to artificially flatten
%the target distribution along $z$\cite{mezei1987, berg1992, lee1993,
%wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}.
%%
%This alleviates the above problem of local trapping
%by forcing the system to travel more
%deliberately along $z$,
%which often represents
%a direction of slow motion.
%%
%To achieve a flat distribution, however,
%the method must be able to construct a
%bias potential that cancels the PMF.
